\vskip 8ex
\centerline{\bf ABOUT \D}
\vskip 14ex

\noindent \D is a general purpose parallel molecular dynamics
simulation package developed at Daresbury Laboratory by I.T.~Todorov
\& W.~Smith.  The DL\_POLY project was developed under the
auspices of the Engineering and Physical Sciences Research Council
(EPSRC) for the EPSRC's Collaborative Computational Project for the
Computer Simulation of Condensed Phases (CCP5), the Computational
Chemistry \& Application Performance Engineering Groups (formerly
Molecular Simulation \& Advanced Research Computing Groups) at
Daresbury Laboratory and the Natural Environment Research Council
(NERC) for the NERC's {\em e}Science project {\em Computational
Chemistry in the Environment} ({\em e}Minerals), directed by M.T.~Dove. \\
\\
\noindent \D is the property of Daresbury Laboratory and is issued
free {\bf under \hyperref[licence]{licence}} to academic institutions pursuing
scientific research of a non-commercial nature.  Commercial
organisations may be permitted a licence to use the package after
negotiation with the owners.  Daresbury Laboratory is the sole
centre for distribution of the package.  Under no account is it to
be redistributed to third parties without consent of the owners. \\
\\
\noindent The purpose of the \D package is to provide software for
academic research that is inexpensive, accessible and free of
commercial considerations.  Users have direct access to source code
for modification and inspection.  In the spirit of the enterprise,
contributions in the form of working code are welcome, provided
the code is compatible with \D in regard to its interfaces and
programming style and it is adequately documented.

\clearpage
