\section{Test Cases}

Because of the size of the data files for the \D
standard test cases, they are not shipped in the standard
download of the \D source.  Instead users are requested
to download them from the CCP5 FTP server as follows:

\begin{lstlisting}
FTP site : ftp.dl.ac.uk
Username : anonymous
Password : your email address
Directory: ccp5/DL_POLY/DL_POLY_4.0/DATA
Files    : test_X.tar.gz
\end{lstlisting}

where `\_X' stands for the test case number.

Remember to use the BINARY data option when transferring these
files.

Unpack the files in the `data' subdirectory using `tar -xzfv'
to create the `TEST\_X' directory.

These are provided so that you may check that your version of
\D is working correctly.  All the jobs are of a size suitable
to test the code in parallel execution.  They not not be
suitable for a single processor computer.  The files are stored
in compressed format.  The test cases can be run by typing

{\sl select } n

\noindent from the {\em execute} directory, where n is the number of
the test case.  The {\sl select} macro will copy the appropriate
CONTROL, CONFIG, and FIELD files to the {\em execute} directory
ready for execution.  The output file OUTPUT may be compared with
the file supplied in the {\em data} directory.

It should be noted that the potentials and the simulation
conditions used in the following test cases are chosen to
demonstrate a limited set of relevant functionality over
a limited extent of molecular systems' complexity only.
{\bf They are not necessarily appropriate for serious
simulation of the test systems.}  In other words, the
tests are not warranted to have well defined force field
in terms of applicability, transferability and fullness
as well as to have a well defined state point - thus
initial configurations may be away from equilibrium, if
physical at all.

\subsection{Test Case 1: Sodium Chloride}

This is a 27,000 ion system with unit electric charges on sodium and
chlorine.  Simulation at 500~K with a NVT Berendsen ensemble.  The SPME
method is used to calculate the Coulombic interactions.

\subsection{Test Case 2: DPMC in Water}

The system consists of 200 DMPC molecules in 9379 and water molecules.
Simulation at 300~K using NVE ensemble with SPME and RATTLE algorithm
for the constrained motion.  The total system size is 51737 atoms.

\subsection{Test Case 3: KNaSi$_{2}$O$_{5}$}

Potassium Sodium disilicate glass (NaKSi$_{2}$O$_{5}$) using two and
three-body potentials.  Some of the two-body potentials are read
from the TABLE file.  Simulation at 1000~K using NVT Nos\'e-Hoover
ensemble with SPME.  Cubic periodic boundaries are in use.  The
total system size is 69120 ions.

\subsection{Test Case 4: Gramicidin A molecules in Water}

The system consists of 8 gramicidin A molecules in aqueous solution
(32,096 water molecules) with total of 99,120 of atoms.  Simulation
at 300~K using NPT Berendsen ensemble with SPME and SHAKE/RATTLE
algorithm for the constrained motion.

\subsection{Test Case 5: SiC with Tersoff Potentials}

The system consists of 74,088 atoms.  Simulation at 300~K using NPT
Nos\'e-Hoover ensemble with Tersoff forces and no electrostatics.

\subsection{Test Case 6: Cu$_{3}$Au alloy with Sutton-Chen (metal) Potentials}

The systems consists of 32,000 atoms.  Simulation at 300~K using NVT
Nos\'e-Hoover ensemble with Sutton-Chen forces and no electrostatics.

\subsection{Test Case 7: Lipid Bilayer in Water}

The systems consists of 12,428 atoms.  Simulation at 300~K using NVT
Berendsen ensemble with SPME and SHAKE/RATTLE algorithm for the constrained motion.

\subsection{Test Case 8 and 9: MgO with Adiabatic and with Relaxed Shell Model}

The system consists of 8,000 (4,000 shells) charged points.  Simulation
at 3000~K using NPT Berendsen ensemble with SPME.

\subsection{Test Case 10: Potential of Mean Force on K+ in Water}

The system consists of 13,500 (500 PMFs) atoms.  Simulation at 300~K
using NPT Berendsen ensemble with SPME and SHAKE/RATTLE algorithm for
the constrained motion.

\subsection{Test Case 11: Cu$_{3}$Au Alloy with Gupta (metal) Potentials}

The system consists of 32,000 atoms.  Simulation at 300~K using NVT
Nos\'e-Hoover ensemble with Gupta forces and no electrostatics.

\subsection{Test Case 12: Cu with EAM (metal) Potentials}

The system consists of 32,000 atoms.  Simulation at 300~K using NPT
Berendsen ensemble with EAM tabulated forces and no electrostatics.

\subsection{Test Case 13 and 14: Al with Analytic and with EAM Tabulated Sutton-Chen (metal) Potentials}

The system consists of 32,000 atoms.  Simulation at 300~K using NVT
Evans ensemble with Sutton-Chen forces and no electrostatics.

\subsection{Test Case 15: NiAl Alloy with EAM (metal) Potentials}

The system consists of 27,648 atoms.  Simulation at 300~K using NVT
Evans ensemble with EAM tabulated forces and no electrostatics.

\subsection{Test Case 16: Fe with Finnis-Sincair (metal) Potentials}

The system consists of 31,250 atoms.  Simulation at 300~K using NPT
Berendsen ensemble with Finnis-Sinclair forces and no electrostatics.

\subsection{Test Case 17: Ni with EAM (metal) Potentials}

The system consists of 32,000 atoms.  Simulation at 300~K using NPT
Berendsen ensemble with EAM tabulated forces and no electrostatics.

\subsection{Test Case 18 and 19: SPC IceVII Water with CBs and with RBs}

The system consists of 11,664 (34,992 atoms) water molecules. Simulation
at 25~K using NVE ensemble with CGM force minimisation and SPME electrostatics.

\subsection{Test Case 20: NaCl Molecules in SPC Water Represented as CBs+RBs}

The system consists of 64 NaCl ion pairs with 4,480 water molecules
represented by constraint bonds and 4,416 water molecules represented
by ridig bodies.  Totalling 26,816 atoms.  Simulation at 295~K using NPT
Berendsen ensemble with CGM energy minimisation and SPME electrostatics.

\subsection{Test Case 21: TIP4P Water: RBs with a Massless Charged Site}

The system consists of 7,263 TIP4P rigid body water molecules totaling
29,052 particles.  Simulation at 295~K using NPT Berendsen ensemble with
CGM energy minimisation and SPME electrostatics.

\subsection{Test Case 22: Ionic Liquid Dimethylimidazolium Chloride}

The system consists of 44,352 ions.  Simulation at 400~K using NPT
Berendsen ensemble, using both particle and rigid body dynamics with SPME electrostatics.

\subsection{Test Case 23: Calcite Nano-Particles in TIP3P Water}

In this case 600 molecules of calcium carbonate in the calcite
structure form 8 nano-particles which are suspended in 6,904 water
molecules, represented by a flexible 3-centre TIP3P model.  Simulation
with SPME electrostatics at 310~K and 1~atmosphere maintained in
a Hoover NPT ensemble.  The system consists of 23,712 ions.

\subsection{Test Case 24: Iron/Carbon Alloy with EEAM}

In this case a steel alloy of iron and carbon in ratio
35132 to 1651 is modelled using an EEAM potential forcefield.
Simulation at 1000~K and 0~atmosphere is maintained in a
Berendsen NPT ensemble.  The system consists of 36,803 particles.

\subsection{Test Case 25: Iron/Cromium Alloy with 2BEAM}

In this case a steel alloy of iron and chromium in ratio
27635 to 4365 is modelled using an 2BEAM potential forcefield.
Simulation at 300~K and 0~atmosphere is maintained in an Evans NVT
isokinetic ensemble.  The system consists of 32,000 particles.

\subsection{Test Case 26 and 27: Hexane and Methanol Melts,
Full Atomistic and Coarse-Grained Simulations}

TEST26 and TEST27 contain a Hexane and a Methanol melt respectively,
(1000 molecules each) modelled by the OPLSAA force-field (FF).
Each system is also supplied in a CG-mapped representation as converted
by VOTCA, \href{http://www.votca.org/}{http://www.votca.org/}\index{WWW},
or DL\_CGMAP \href{http://www.ccp5.ac.uk/projects/ccp5\_cg.shtml}{http://www.ccp5.ac.uk/projects/ccp5\_cg.shtml}\index{WWW}.

These test cases are to exemplify the Coarse-Graining (CG) procedure
(see Chapter~\ref{coarse-graining}), including FA-to-CG mapping and
obtaining the PMF data by means of Boltzmann Inversion \cite{reith-03a}.
As a result, \D could be used for simulating a CG system with numerically
defined, tabulated FFs, see TABBND, TABANG, TABDIH and TABINV files for
intra-molecular potentials, and TABLE for inter-molecular (short-range, VDW)
potentials.

Both tests are also available as parts of the tutorial cases from the
VOTCA package \cite{ruhle-09a}.  Therefore, the CONFIG, CONTROL and FIELD
input files are fully consistent with the corresponding setup files found
in the VOTCA tutorial directories ``csg-tutorials/hexane'' and ``csg-tutorials/methanol'.

\subsection{Test Case 28: Free Energy calculation for Buthane with PLUMED}

Free Energy calculation for Buthane with respect to the dihedral angle as
collective variable.  We use umbrella sampling as implemented in PLUMED.

PLUMED enabling in CONTROL:
\begin{lstlisting}
plumed input umbrella.dat
\end{lstlisting}

Contents of umbrella.dat:
\begin{lstlisting}
phi: TORSION ATOMS=1,2,3,4
restraint-phi: RESTRAINT ARG=phi KAPPA=500 AT=1.20
PRINT STRIDE=10 ARG=phi,restraint-phi.bias FILE=COLVAR
\end{lstlisting}

Two extra output files are generated in this case: OUTPUT.PLUMED and COLVAR.

{\bf Note}, a \D version with PLUMED enabled is used for this.

\section{Benchmark Cases}

\D benchmark test cases are available to download them from the CCP5 FTP
server as follows:

\begin{lstlisting}
FTP site : ftp.dl.ac.uk
Username : anonymous
Password : your email address
Directory: ccp5/DL_POLY/DL_POLY_4.0/BENCH
\end{lstlisting}

The \D authors provide these on an "AS IS" terms.  For more information
refer to the README.txt file within.
